---
title: "Homework 2"
author: Alexey Abramov
output: 
  github_document
---

## Setup
```{r, setup}
library(tidyverse)
library(readxl)
library(chron)
```

# Problem 1
 
### Mr. Trash Wheel, sports balls data.

#### Read in and clean up the Mr. Trashwheel dataset
```{r}
trashwheel_df = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
      janitor::clean_names() %>% 
      drop_na(dumpster) %>% 
      mutate(
        sports_balls = round(sports_balls),
        sports_balls = as.integer(sports_balls))
```

### Trash Wheel, precipitation data.

#### Read in precipitation excel spreadsheet for the 2018 and 2017 data.  

```{r}
precip_2018 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2018 Precipitation",
      skip = 1) %>% 
      janitor::clean_names() %>% 
      drop_na(month) %>% 
      mutate(year = 2018) %>% 
      relocate(year)

precip_2017 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2017 Precipitation",
      skip = 1) %>% 
      janitor::clean_names() %>% 
      drop_na(month) %>% 
      mutate(year = 2017) %>% 
      relocate(year)
```

#### Binding 2017 and 2018 precipitation data.

```{r}
  precip_df = bind_rows(precip_2017, precip_2018)
```

Creating a month number to month name tibble:

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )
```

#### Cleaning up 2017 and 2018 Precipitation Data 

Including a month name column and relocating columns for improved visualization.
```{r}
precip_df_2 = 
  left_join(precip_df, month_df, by = "month") %>% 
  relocate(year, month_name, month, total)
```

Total precipitation in 2018 calculation

```{r}
sum_precip_2018 = select(precip_2018, total) %>% 
  sum()
```

Median number of sports balls in 2017 calculation

```{r}
trashwheel_df_2017_sports_balls = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
      janitor::clean_names() %>% 
      drop_na(dumpster) %>% 
      filter(year == "2017") %>% 
      pull(sports_balls) %>% 
      median()
```


### Mr. Trash Wheel Discussion

These data are obtained from the Mr. Trash Wheel dataset which reports on discarded items collected from the Inner Harbor in Baltimore, MD and stores it in a dumpster. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. These data also include precipitation amounts by month.  

In this data set, the median number of sports balls collected in 2017 is `r trashwheel_df_2017_sports_balls` and the sum precipitation in 2017 was `r sum_precip_2018` inches.

# Problem 2

## NYC Transit dataset

#### Read in dataset, cleaning up column names

```{r}
nyc_transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entrance_type, entry, vending, ada)
```

#### Data cleaning, and recoding entry column to logical vector

```{r}
nyc_transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
   mutate(entry = ifelse(entry == "NO",0,1))
```

## Discussion 

This dataset contains information regarding the New York City transit subway system and reports station name and GPS coordinates, routes served, as well as entrance information highlighting ADA compliance and vending.  The dataset has `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)` columns.     

#### Calculating the number of distinct stations 

There are `r nrow(distinct(nyc_transit_df, station_name, line))` stations in the NYC transit system.  

```{r}
nrow(distinct(nyc_transit_df, line, station_name))
```

#### Calculating the number of ADA compliant stations

```{r}
ada_compliant = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
   mutate(ada = ifelse(ada == "TRUE",1,0)) %>% 
    filter(ada == "1") %>% 
    distinct(line, station_name)

nrow(ada_compliant)
```

There are `r nrow(ada_compliant)` ADA compliant stations in the NYC transit system.

#### Calculating the number of stations with entry that do not have vending.

```{r}
entry_vending_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>%
     select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
        mutate(vending = ifelse(vending == "YES",1,0)) %>% 
        mutate(entry = ifelse(entry == "NO",0,1)) %>% 
          filter(vending == "0") %>% 
          filter(entry == "1") %>% 
            distinct(line, station_name)

nrow(entry_vending_df)
```

There are `r nrow(entry_vending_df)` stations that have entrances but do not have vending.  There are a total of `r nrow(distinct(nyc_transit_df, line, station_name))` stations, which means `r nrow(entry_vending_df)/nrow(distinct(nyc_transit_df, line, station_name))` do not have vending.  

#### Reformatting to create distinct route name and route number variables. 

```{r}
route_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>%
     select(line, station_name, route1:route11, entrance_type, entry, 
            vending, ada) %>% 
              mutate_at(vars(route1:route11), as.character) %>% 
                  pivot_longer(route1:route11, 
                  names_to = "route_number", 
                  values_to = "route_name",
                  names_prefix = "route") %>% 
                    drop_na("route_name") %>% 
                      distinct_at(vars(line, station_name, route_name, 
                                       route_number))
```

```{r}

```


#### Calculating the number of distinct A train stations

```{r}
route_A_df = filter(route_df, route_name == "A")
nrow(route_A_df)

ada_compliant_A_df= 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>%
     select(line, station_name, route1:route11, ada) %>% 
              mutate_at(vars(route1:route11), as.character) %>% 
                  pivot_longer(route1:route11, 
                  names_to = "route_number", 
                  values_to = "route_name",
                  names_prefix = "route") %>% 
                    drop_na("route_name") %>% 
                      distinct(station_name, line, ada)

nrow(ada_compliant_A_df)
```

There are `r nrow(route_A_df)` distinct A stations in the NYC transit system. Of which, `r nrow(ada_compliant_A_df)` are ADA compliant. 

# Problem 3

#### Read in FiveThirtyEight data pols-month

Data cleaning pols_month dataframe.
```{r}
library(lubridate)

pols_month_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
    janitor::clean_names() %>% 
      mutate(mon = ymd(mon)) %>% 
      mutate_at(vars(mon), funs(year, month, day)) %>% 
      mutate(president = 
               case_when(
                  prez_gop == 1 ~ "gop",
                  prez_dem == 1 ~ "dem")) %>% 
      subset(select = -c(prez_gop, prez_dem, day, mon))
```

Reformatting month number with the month name.

```{r}
pols_month_df = 
  left_join(pols_month_df, month_df, by = "month")  %>%        
  relocate(year, month_name, president)
```

Data cleaning snp dataframe.

```{r}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
    janitor::clean_names() %>% 
    separate(date, into = c("month", "day", "year"), sep="/")
```

Binding to reformat month number to month name.
```{r}
snp_df = 
  left_join(snp_df, month_df, by = "month")
```

Problem 3
This problem uses the FiveThirtyEight data; these data were gathered to create the interactive graphic on this page. In particular, we’ll use the data in pols-month.csv, unemployment.csv, and snp.csv. Our goal is to merge these into a single data frame using year and month as keys across datasets.

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

Join the datasets by merging snp into pols, and merging unemployment into the result.

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

Note: we could have used a date variable as a key instead of creating year and month keys; doing so would help with some kinds of plotting, and be a more accurate representation of the data. Date formats are tricky, though. For more information check out the lubridate package in the tidyverse.

Optional post-assignment survey
If you’d like, a you can complete this short survey after you’ve finished the assignment.