---
title: "Homework 2"
author: Alexey Abramov
output: 
  github_document
---

## Setup
```{r, setup}
library(tidyverse)
library(readxl)
```

# Problem 1
 
### Mr. Trash Wheel, sports balls data.

#### Read in and clean up the Mr. Trashwheel dataset
```{r}
trashwheel_df = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
      janitor::clean_names() %>% 
      drop_na(dumpster) %>% 
      mutate(
        sports_balls = round(sports_balls),
        sports_balls = as.integer(sports_balls))
```

### Trash Wheel, precipitation data.

#### Read in precipitation excel spreadsheet for the 2018 and 2017 data.  

```{r}
precip_2018 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2018 Precipitation",
      skip = 1) %>% 
      janitor::clean_names() %>% 
      drop_na(month) %>% 
      mutate(year = 2018) %>% 
      relocate(year)

precip_2017 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2017 Precipitation",
      skip = 1) %>% 
      janitor::clean_names() %>% 
      drop_na(month) %>% 
      mutate(year = 2017) %>% 
      relocate(year)
```

#### Binding 2017 and 2018 precipitation data.

```{r}
  precip_df = bind_rows(precip_2017, precip_2018)
```

Creating a month number to month name tibble:

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )
```

#### Cleaning up 2017 and 2018 Precipitation Data 

Including a month name column and relocating columns for improved visualization.
```{r}
precip_df_2 = 
  left_join(precip_df, month_df, by = "month") %>% 
  relocate(year, month_name, month, total)
```

Total precipitation in 2018 calculation

```{r}
sum_precip_2018 = select(precip_2018, total) %>% 
  sum()
```

Median number of sports balls in 2017 calculation

```{r}
trashwheel_df_2017_sports_balls = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
      janitor::clean_names() %>% 
      drop_na(dumpster) %>% 
      filter(year == "2017") %>% 
      pull(sports_balls) %>% 
      median()
```


### Mr. Trash Wheel Discussion

These data are obtained from the Mr. Trash Wheel dataset which reports on discarded items collected from the Inner Harbor in Baltimore, MD and stores it in a dumpster. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. These data also include precipitation amounts by month.  

In this data set, the median number of sports balls collected in 2017 is `r trashwheel_df_2017_sports_balls` and the sum precipitation in 2017 was `r sum_precip_2018` inches.

# Problem 2

## NYC Transit dataset

#### Read in dataset, cleaning up column names

```{r}
nyc_transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entrance_type, entry, vending, ada)
```

#### Data cleaning, and recoding entry column to logical vector

```{r}
nyc_transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
   mutate(entry = ifelse(entry == "NO",0,1))
```

## Discussion 

This dataset contains information regarding the New York City transit subway system and reports station name and GPS coordinates, routes served, as well as entrance information highlighting ADA compliance and vending.  The dataset has `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)` columns.     

#### Calculating the number of distinct stations 

There are `r nrow(distinct(nyc_transit_df, station_name, line))` stations in the NYC transit system.  

```{r}
nrow(distinct(nyc_transit_df, line, station_name))
```

#### Calculating the number of ADA compliant stations

```{r}
ada_compliant = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
   mutate(ada = ifelse(ada == "TRUE",1,0)) %>% 
    filter(ada == "1") %>% 
    distinct(line, station_name)

nrow(ada_compliant)
```

There are `r nrow(ada_compliant)` ADA compliant stations in the NYC transit system.

#### Calculating the number of stations with entry that do not have vending.

```{r}
entry_vending_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>%
     select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
        mutate(vending = ifelse(vending == "YES",1,0)) %>% 
        mutate(entry = ifelse(entry == "NO",0,1)) %>% 
          filter(vending == "0") %>% 
          filter(entry == "1") %>% 
            distinct(line, station_name)

nrow(entry_vending_df)
```

There are `r nrow(entry_vending_df)` stations that have entrances but do not have vending.  There are a total of `r nrow(distinct(nyc_transit_df, line, station_name))` stations.  

#### Reformatting to create distinct route name and route number variables. 

```{r}
route_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>%
     select(line, station_name, route1:route11, entrance_type, entry, 
            vending, ada) %>% 
              mutate_at(vars(route1:route11), as.character) %>% 
                  pivot_longer(route1:route11, 
                  names_to = "route_number", 
                  values_to = "route_name",
                  names_prefix = "route") %>% 
                    drop_na("route_name") %>% 
                      distinct_at(vars(line, station_name, route_name, 
                                       route_number))
```

```{r}

```


#### Calculating the number of distinct A train stations

```{r}
route_A_df = filter(route_df, route_name == "A")
nrow(route_A_df)

ada_compliant_A_df= 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>%
     select(line, station_name, route1:route11, ada) %>% 
              mutate_at(vars(route1:route11), as.character) %>% 
                  pivot_longer(route1:route11, 
                  names_to = "route_number", 
                  values_to = "route_name",
                  names_prefix = "route") %>% 
                    drop_na("route_name") %>% 
                      distinct(station_name, line, ada)

nrow(ada_compliant_A_df)
```

There are `r nrow(route_A_df)` distinct A stations in the NYC transit system. Of which, `r nrow(ada_compliant_A_df)` are ADA compliant. 

# Problem 3

#### Read in FiveThirtyEight data pols-month

```{r}
library(lubridate)

pols_month_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
    janitor::clean_names() %>% 
      mutate(mon = ymd(mon)) %>% 
      mutate_at(vars(mon), funs(year, month, day)) %>% 
      relocate(year, month, day)
```

