---
title: "Homework 2"
output: github_document
---

# Setup

```{r, setup}
library(tidyverse)
library(readxl)
```

## Problem 1
 
### Trash Wheel, sports balls data.

Read in and clean up the Mr. Trashwheel dataset
```{r}
trashwheel_df = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
      janitor::clean_names() %>% 
      drop_na(dumpster) %>% 
      mutate(
        sports_balls = round(sports_balls),
        sports_balls = as.integer(sports_balls))
```

### Trash Wheel, precipitation data.

Read in precipitation excel spreadsheet for the 2018 and 2017 data.  

```{r}
precip_2018 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2018 Precipitation",
      skip = 1) %>% 
      janitor::clean_names() %>% 
      drop_na(month) %>% 
      mutate(year = 2018) %>% 
      relocate(year)

precip_2017 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2017 Precipitation",
      skip = 1) %>% 
      janitor::clean_names() %>% 
      drop_na(month) %>% 
      mutate(year = 2017) %>% 
      relocate(year)
```

### Binding 2017 and 2018 precipitation data.

```{r}
  precip_df = bind_rows(precip_2017, precip_2018)
```

Creating a month number to month name tibble:

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )
```

### 2017 and 2018 Precipitation Data 

Include a month name column and relocating columns for improved visualization.
```{r}
precip_df_2 = 
  left_join(precip_df, month_df, by = "month") %>% 
  relocate(year, month_name, month, total)
```

Total precipitation in 2018

```{r}
sum_precip_2018 = select(precip_2018, total) %>% 
  sum()
```

Median number of sports balls in 2017

```{r}
trashwheel_df_2017_sports_balls = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
      janitor::clean_names() %>% 
      drop_na(dumpster) %>% 
      filter(year == "2017") %>% 
      pull(sports_balls) %>% 
      median()
```


### Discussion

These data are obtained from the Mr. Trash Wheel dataset which reports on discarded items collected from the Inner Harbor in Baltimore, MD and stores it in a dumpster. There are a total of `r nrow(trashwheel_df)`in our final dataset. These data also include precipitation amounts by month.  

In this data set, the median number of sports balls collected in 2017 is `r trashwheel_df_2017_sports_balls` and the sum precipitation in 2017 was `r sum_precip_2018` inches.

## Problem 2

### Read in dataset

```{r}
nyc_transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv" %>% 
             janitor::make_clean_names() %>% 
             drop_na(dumpster)
  )
```

